Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-25 paper/main.tex
edits through pdf line 364, source line 686

That the knowledge predictions derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that the text embedding space bears at
least some relationship to participants' knowledge. But what does that
relationship look like as we move through the embedding space? For example,
suppose we know that a participant answers a question (at embedding coordinate
$X$) correctly. As we move away from $X$ in the embedding space, how does quiz
performance ``fall off'' with distance? Or, suppose the participant instead
answered that same question \textit{in}correctly. Again, as we move away from
$X$ in the embedding space, how do the chances that the participant does
\textit{not} know about the content change with distance? We reasoned that,
assuming our space is capturing something about how participants actually
organize their knowledge, conceptual knowledge right around $X$ should be
similar to the participant's knowledge of the content at $X$. And at another
extreme, at some distance (after moving sufficiently far away from $X$), our
guesses about what participants know (based on their response to the question
at location $X$) should be no better than guessing based on their overall
proportion of correctly answered questions—i.e., if $Y$ is very far away from
$X$, all we can do with the participant's response to $X$ is guess that ``their
performance on quiz questions about $Y$ is about equal to their average
performance on quiz questions about any material.''

>>>>

That the knowledge predictions derived from the text embedding space reliably
distinguish between correctly and incorrectly answered held-out questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships within this
 space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
correctly answered a question at embedding coordinate $x$. As we move farther away
from $x$ in the embedding space, how does the likelihood that the participant knows about the content at a given location ``fall off'' with distance?
Conversely, suppose the participant instead answered that same question
\textit{in}correctly. Again, as we move farther away from $x$ in the embedding space,
how does the likelihood that the participant does \textit{not} know about a coordinate's
content change with distance? We reasoned that, assuming our embedding space is capturing
something about how individuals actually organize 
their knowledge, a participant's ability to answer questions 
embedded very close to $x$ should tend to be similar to their 
ability to answer the question embedded \textit{at} $x$. Whereas 
at another extreme, once we reach some sufficiently large 
distance from $x$, our ability to infer whether or not a participant 
will correctly answer a question based on their ability to answer 
the question at $x$ should be no better than guessing based on 
their \textit{overall} proportion of correctly answered questions. In 
other words, beyond the maximum distance at which the participant's
 ability to answer the question at $x$ is informative of their ability to 
 answer a second question at location $y$, then guessing the outcome 
 at $y$ based on $x$ should be no more successful than guessing 
 based on a measure that does not consider embedding space distance.
